---
title: 'Activity: Popular Names'
author: "name"
date: "June 3, 2024"
output:
  html_document:
    df_print: paged
  html_notebook: default
  pdf_document: default
---


# Grading

Assignment is worth a total of 10 points

* [1 point] Turn in BOTH RNotebook (.nb.html) and the RMarkdown (.Rmd) file to Canvas
* [1 point] Changes pushed to GitHub
* [1 point] Step 1: show how you used summary functions to explore the raw data set
* [1 point] Step 3 and 4: attempt to answer all questions about the data
* [1 point] Step 5: explanation of data wrangling steps in English (not code)
* [3 points] Step 7: Data wrangling
    * correct use of `filter()`
    * correct use of `group_by()`
    * correct use of `summarise()`
* [2 points] `ggplot()` graph of first names 



# Set Up

```{r, message = F, warning = F}
# load necessary packages
library(tidyverse)
library(mosaic)
library(esquisse)

devtools::install_github("mdbeckman/dcData")
library(dcData)

# Load the BabyNames data into RStudio environment (from dcData package)
data("BabyNames", package = "dcData")
```

# Activity

**Goal**: By the end of the activity, each team will create a graph (like Figure A.1 on p 159) showing ups and downs in the popularity of the names at your table according to the U.S. social security registry.  

For steps 5-7, pick any FIVE names you wish.

I choose Adam , John , Conner , Gunner , Steve  


### Step 1
```{r}
data("BabyNames")

str(BabyNames)
summary(BabyNames)

##### What are the variables are available?

# The variables are : Name , sex , count , year 

##### What is a case?

# Case is a single observation , which in this case is specific name given to the babies of a specific sex in a specific year 

```


## Look at the data a few different ways (see Ch 9 for more).
```{r}
# We can look at the first few rows of the data 

head(BabyNames)

```

### Step 2

[This should be done by hand on paper, nothing to show in the HTML doc]


### Step 3 

##### What variable(s) from the raw data table do not appear at all in the graph?

The variable sex does not appear in the graph.

##### What variable(s) in the graph are similar to corresponding variables in the raw data table, but might have been transformed in some way?

The variables year and name in the graph correspond directly to the raw data. The Count variable is transformed to total to represent the total number of births per year for each name.

### Step 4

##### Have cases been filtered out?

Yes, the cases have been filtered to include only the selected five names.

##### Have cases been grouped and summarized within groups in any way?

Yes, the cases have been grouped by year and name, and then summarized to get the total number of births for each name per year.

##### Have any new variables been introduced?  If so, what the relationship between the new variables and existing variables?

Yes, the new variable total has been introduced, which represents the total number of births for each name per year. This variable is derived from the n variable by summing up the counts for each name per year.


### Step 5

Using the BabyNames data set, filter the data to display just the 5 names that you picked.  Then group the cases by year and name, and count how many total births per year were recorded for each of the names (regardless of sex).  

1. Select the five names

2. Filter the BabyNames dataset for only those 5 names selected in Step-1 and store the output data in separate variable as a dataframe

3. Use "group_by" function to group by the output data from Step-2 at "name" and "year" level

4. Use "summarise" function to get the total number of births  recorded for each of the names per year on the grouped by data from Step-3


### Step 6

[This should be done by hand on paper, nothing to show in the HTML doc]

### Step 7
#### Create my glyph ready data set here
```{r}
# Define the selected names
selected_names <- c("Adam" , "John" , "Conner" , "Gunner" , "Steve")

# Filter the data for the selected names
filtered_data <- BabyNames %>%
  filter(name %in% selected_names)

# Group and summarize the data
glyph_ready_data <- filtered_data %>%
  group_by(year, name) %>%
  summarize(total = base::sum(count))

# Arrange the data for plotting
glyph_ready_data <- glyph_ready_data %>%
  arrange(year, name)

# Display the glyph-ready data
head(glyph_ready_data)


```

### Graph It!

Here's the sample code provided by the book:

```
GlyphReadyForm %>%
  ggplot(aes(x = year, y = total, group = name)) + 
  geom_line(size = 1, alpha = 0.5, aes(color = name)) + 
  ylab("Popularity") + 
  xlab("Year") +
  ggtitle("Popularity of Names Over Time")
```

If you prepared your `GlyphReadyForm` data set correctly, you should just be able to copy and paste this code. Alternatively, you can use `mplot` or `esquisse` to create your graph. If you choose to use these interactive functions remember, do not include the `mplot` or `esquisse` functions in your RMarkdown document. Instead, use these functions in the console and copy the generated code into your RMarkkdown document. In addition, we want lines, not dots.




